---
title: "PrismLows"
author: "A Hessl"
date: "1/3/2019"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Daily PRISM Data

Read and extract annual winter lows from PRISM Data
```{r install libraries, message=TRUE, warning=TRUE}
rm(list=ls())
library(tidyverse)
library(lubridate)
```

Read in all prism files in data/daily_cv directory, reformat and arrange as a list
```{r read.prism}
#call function read.prism.R
source('Functions/read.prism.R')
pgrid.list <- read.prism("daily_pr")
```

Select the data from read.prism list, derive min monthly T, and add time variable for plotting

```{r min monT}
#Select just the mean of the dataframes to work with
pgrid <- pgrid.list$pgrid.df

## Minimum monthly temperatures
mon.min <- pgrid %>% group_by(year=year(pdate), month=month(pdate)) %>% summarize(m.min=min(tmin))

#add readable time variable for plotting
mon.min$pdate <- paste0(mon.min$year, "-", mon.min$month, "-1")
```

```{r plot min monT}
plot(as.Date(mon.min$pdate), mon.min$m.min, type='l', xlab="Year", ylab="MinMonT(C)", main="Min Monthly T")
```

Count of Degree Days Below a Threshold
```{r threshold}
#Find duration of the low below a critical threshold
thresh <- 0
pgrid$thresh <- pgrid$tmax < thresh

pgrid$year <- year(pgrid$pdate)
pgrid$month <- month(pgrid$pdate)
pgrid$day <- day(pgrid$pdate)

#Adjust Nov and Dec to be subsequent year
prev.win <- c(11, 12)
adj.year <- ifelse (pgrid$month %in% prev.win, pgrid$year+1, pgrid$year)
pgrid$year <- adj.year

thresh.event <- pgrid[pgrid$thresh %in% 'TRUE', -c(1:5)] 

#Count number of days tmax < thresh
thresh.count <- aggregate(thresh.event, by=list(thresh.event$year), sum)[,1:2]
names(thresh.count) <- c("Year", "Count")
thresh.count <- thresh.count[-38,]

#barplot(thresh.count$Count, names=thresh.count$Year, ylab="Count (neg degree days)", xlab="Year")

```

```{r lolliplot}
#remake as lollipop charts in GGplot
library(ggplot2)
theme_set(theme_bw())

# Plot
ggplot(thresh.count, aes(x=Year, y=Count)) + 
  geom_point(size=2) + 
  geom_segment(aes(x=Year, 
                   xend=Year, 
                   y=0, 
                   yend=Count)) + 
  labs(title="Lollipop Chart", 
       subtitle="Count Negative Degree Days", 
       caption="source: Prism Daily") + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6)) +
  scale_x_discrete (limits=c(1981:2017))
```

Cumulative Sum of Negative Degree Days
```{r negative degree days}
#Select only events below thresh and only needed cols
neg.days <- pgrid[pgrid$thresh %in% 'TRUE', c("tmax", "year", "month")] 

#cumulative sum of negative degree days
neg.dd <- aggregate(neg.days, by=list(neg.days$year), sum)[,c(1,2)]
names(neg.dd) <- c("Year", "sumTmax")
neg.dd <- neg.dd[-38,]

#barplot(neg.dd$sumTmax, names=neg.dd$Year, ylim=c(0,-350), ylab="Sum (neg degree days)", xlab="Year")
```

```{r lollipop DD sum}
# Plot
ggplot(neg.dd, aes(x=Year, y=sumTmax)) + 
  geom_point(size=2) + 
  geom_segment(aes(x=Year, 
                   xend=Year, 
                   y=0, 
                   yend=sumTmax)) + 
  labs(title="Lollipop Chart", 
       subtitle="Sum Negative Degree Days", 
       caption="source: Prism Daily") + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6)) +
  scale_x_discrete (limits=c(1981:2017))
```

Duration of Events
```{r duration}
runs <- rle(pgrid$thresh)

pgrid$runs <- rep(runs$lengths, runs$lengths) #add a new vector that tells the lengths and positions of the runs

#give each run a unique identifier to aggregate on
events <- length(runs$values)
pgrid$event.id <- rep(1:events, c(runs$lengths))
```


Events below threshold (TRUE) & longer than 5 days are kept, group by year and month
```{r events plot}
long.event <- pgrid[pgrid$thresh %in% 'TRUE' & pgrid$runs >= 5, -c(1:6)] 

#start date and length of each run below threshold
long.st.date <- aggregate(long.event, by=list(long.event$event.id), min)

num.long.yr <- aggregate(long.st.date, by=list(long.st.date$year), length)[, c(1, 6)]

#barplot(num.long.yr$runs[-37], names=num.long.yr$Group.1[-37], ylab="No. Long Freeze (>5 days)", xlab="Year")

```

```{r lollipop long event}
# Plot
ggplot(num.long.yr[-36,], aes(x=Group.1, y=runs)) + 
  geom_point(size=2) + 
  geom_segment(aes(x=Group.1, 
                   xend=Group.1, 
                   y=0, 
                   yend=runs)) + 
  labs(title="Lollipop Chart", 
       subtitle="Long Cold Runs", 
       caption="source: Prism Daily",
       x="Year",
       y="Num Runs") + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6)) +
  scale_x_discrete (limits=c(1981:2017))
```

### Monthly Data
Read in Monthly Data
```{r plot monthly data}

rm(list=ls())
library(lubridate)
library(tidyverse)
source('Functions/read.prism.R')
pgrid.list <- read.prism("mon_cv")

#Read the averaged data from pgrid.list
pgridm <- pgrid.list$pgrid.df

#add readable time variable for plotting
pgridm$pdate <- paste0(pgridm$pdate, "-01")

pgridm_sub <- pgridm[pgridm$pdate >= "1970-01-01",]
{plot(as.Date(pgridm_sub$pdate), pgridm_sub$tmin, type='l')
abline(h=mean(pgridm_sub$tmin))}

```

Plot Annual Minimum Temperature
```{r plot annual min}
#adjust Nov, Dec to subsequent year
pgridm$year <- year(pgridm$pdate)
pgridm$month <- month(pgridm$pdate)

prev.win <- c(11, 12)
adj.year <- ifelse (pgridm$month %in% prev.win, pgridm$year+1, pgridm$year)
pgridm$adj.year <- adj.year

win.min <- aggregate(pgridm, by=list(pgridm$adj.year), min)[,c(3,7)]
{plot(win.min$year, win.min$tmin, type='l', xlab='Year', ylab='MinMonT', main="Annual Low T")
text(1964, -16, "Ohio R Froze")}
```

Lollipop Plot of tmin and tmax for each winter (N, D, J, F)
 ```{r lollipop tmin tmax}
# # select just winter (11,12, 1, 2)
# winter <- c(11, 12, 1, 2)
# pgridm.win <- pgridm[pgridm$month %in% winter, c(2, 4, 6, 7)]
# aggregate (pgridm.win, by=list())
# 
# ggplot(pgridm) +
#   geom_segment( aes(x=x, xend=x, y=value1, yend=value2), color="grey") +
#   geom_point( aes(x=x, y=value1), color=rgb(0.2,0.7,0.1,0.5), size=3 ) +
#   geom_point( aes(x=x, y=value2), color=rgb(0.7,0.2,0.1,0.5), size=3 ) +
#   coord_flip() 
#  
# # With a bit more style
# ggplot(data) +
#   geom_segment( aes(x=x, xend=x, y=value1, yend=value2), color="grey") +
#   geom_point( aes(x=x, y=value1), color=rgb(0.2,0.7,0.1,0.5), size=3 ) +
#   geom_point( aes(x=x, y=value2), color=rgb(0.7,0.2,0.1,0.5), size=3 ) +
#   coord_flip()+
#   theme_light() +
#   theme(
#     legend.position = "none",
#     panel.border = element_blank(),
#   ) +
#   xlab("") +
#   ylab("Value of Y")
# 
```
